---
title: Machine Learning in R
subtitle: R-Ladies Colombo 
author: Kasun Bandara
date: 29 March, 2021
institute: Melbourne Centre for Data Science, University of Melbourne, Australia.
beameroption: "show notes"
output: 
  binb::metropolis:
    toc: false
    keep_tex: true
    citation_package: natbib
  includes:
    in_header: metropolis/header.tex  
fontsize: 12pt
header-includes:
  - \usepackage{subfig}
---

```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.path='figs/')
knitr::opts_chunk$set(cache.path='cache/')
knitr::opts_chunk$set(
                  fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      }
                  )

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```


\begin{figure}
\includegraphics[scale=0.22]{images/kasun}
\end{figure}

# Introduction

## About me
- 2015 Graduated in Computer Science from University of Colombo School of Computing
- 2015 Joined WSO2 Inc. as a Software Engineer
- 2016-2020 Ph.D. in Computer Science, Monash University, Australia
  - Topic: Forecasting In Big Data With Recurrent Neural Networks
  - Machine Learning for Time Series Forecasting
  - Research Internship at Walmart Labs, San Francisco, USA
  - Research Scientist at Turning Point, Melbourne, Australia
  - Data Science Tutor, Faculty of IT, Monash University
- 2021 Research Fellow, University of Melbourne

## About me (2)
- Research Interests
  - Global Forecasting Models
  - Hierarchical Forecasting
  - Retail sales/demand forecasting
  - Renewable energy production forecasting (solar)
  
- Competition Fanatic
  - M5 Forecasting Competition (**Gold Medalist**)
  - IEEE CIS Energy Forecasting Competition (**4th Place**)
  - Air-Liquide Energy Forecasting Competition (**4th Place**)
  - ANZ Customer Segmentation Challenge (**Top Performer**) 

## What is Data Science ?
Data Science is an interdisciplinary field that permits you to extract information from organized or unstructured data.

\begin{figure}
  \includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{images/datascience.jpeg}
  \caption{An intersection of many fields of science%
    \footnote{%
     \tiny{Image source: https://medium.com/believing-these-8-myths-about-what-is-data-science-keeps-you-from-growing-528f1bd240dc} 
    }%
  }
\end{figure}

## Data Science Life Cycle
Known as the O.S.E.M.N. framework.

\begin{figure}
  \includegraphics[width=.8\textwidth,height=.8\textheight,keepaspectratio]{images/osemn.png}
  \caption{Data Science Process%
    \footnote{%
     \tiny{Image source: https://towardsdatascience.com/5-steps-of-a-data-science-project-lifecycle-26c50372b492} 
    }%
  }
\end{figure}

## Obtain (O)
- Retrieving data from multiple sources of inputs. \
 \begin{itemize}
      \item Structured Data: RDBMS, Tabular Data, CSV, TSV.
      \item Unstructured Data: NoSQL Databases, API Data (Twitter, Facebook).
\end{itemize}
 \vspace{2mm}
- Databases: \texttt{\{odbc\}}
 \vspace{2mm}
- Scraping data from websites: \texttt{\{rvest\}}
 \vspace{2mm}
- Data platforms: **Kaggle**, **UCI**, **Competition Datasets**, **Government APIs**

## Example of \texttt{\{rvest\}}

```{r rvestcode, eval=FALSE, size="tiny"}
library(rvest)
library(dplyr)
set.seed(1234)

# reading the HTML page (Lord of the Rings)
lor_movie <- read_html("https://www.imdb.com/title/tt0120737/")

# Scraping the movie rating.
lor_movie %>%
  html_node("strong span") %>%
  html_text() %>%
  as.numeric()
#[1] 8.8

# Scraping the cast.
lor_movie %>%
  html_nodes("#titleCast .itemprop span") %>%
  html_text()

# Scraping the movie poster.
lor_movie %>%
  html_nodes("#img_primary img") %>%
  html_attr("src")
```  

## Scrub (S)
- Also known as \textbf{data pre-processing}, \textbf{data wrangling}. \
 \vspace{2mm}
- Converting the data into a unified, suitable format
\begin{itemize}
      \item Easier for the data exploration process.
      \item What your predictive algorithm expects ?
      \item \textbf{tidyverse} \texttt{\{dplyr,tidyr,stringr,tibble,purr,ggplot2\}}
\end{itemize}
 \vspace{2mm}
- Handles data issues 
\begin{itemize}
      \item Cleaning: Missing values, Outliers, Noisy data.
      \item Transformation: Normalisation, Feature Discretization.
      \item Reduction: Feature selection, Dimensionality reduction.
\end{itemize}

## Missing Value Imputation

```{r simputation, eval=FALSE, size="tiny"}
library(simputation)
set.seed(1234)

# Loading iris dataset and randomly inserting NAs.
df <- iris
df_NA <- as.data.frame(lapply(df, function(imp) imp[ sample(c(TRUE, NA), 
        prob = c(0.85, 0.15), size = length(imp), replace = TRUE)]))

# Using median to impute the missing values.
median_imputed <- impute_median(df_NA, 
                                Sepal.Length ~ Species)

# Using linear regression to impute the missing values.
linear_imputed <- impute_lm(df_NA, Sepal.Length ~ Sepal.Width + Species)

# Using CART algorithm to impute the missing values.
cart_imputed <- impute_cart(df_NA, Species ~ .)

# Imputing multiple variables at once.
multivariable_imputed <- impute_rlm(df_NA, Sepal.Length + Sepal.Width 
                                    ~ Petal.Length + Species)

# Imputing using a pre-trained model.
model <- lm(Sepal.Length ~ Sepal.Width + Species, data=iris)
model_imputed <- impute(df_NA, Sepal.Length ~ model)

``` 

## Dealing with Outliers
- A data point that differs significantly from other observations. \
 \vspace{2mm}
- Observations that distort your analysis.
\begin{itemize}
      \item Boxplot visualisation: \texttt{\{ggplot2\}}
      \item Grubbs’s test, Dixon’s test, Rosner’s test: \texttt{\{outliers\}}
      \item Outlier detection algorithms: \texttt{\{OutlierDetection\}}
      \item \textbf{outlierTest()} from \texttt{\{car\}}
      \item \textbf{lofactor()} from \texttt{\{DMwR\}} (Local Outlier Factor)
\end{itemize}
 \vspace{2mm}
- Anomaly detection is itself a different research area ! 
\begin{itemize}
      \item One Class SVM, IsolationForest
      \item Unsupervised algorithms (Clustering)
      \item Time series: \texttt{\{tsoutliers,oddstream,stray\}}
\end{itemize}


## Feature Selection
- Removing redundant features from the dataset.
 \vspace{2mm}
- Computational complexity, Address model overfitting.
 \vspace{2mm}
- \textbf{Filter Methods}
\begin{itemize}
      \item Features are selected based on a statistical score.
      \item Independent of any machine learning algorithm.
      \item \textbf{Pearson’s Correlation, Chi-Square, PCA}
\end{itemize}
 \vspace{2mm}
- \textbf{Wrapper Methods}
\begin{itemize}
      \item A subset of features are used to train a model.
      \item Forward, Backward, Recursive elimination.
      \item Inbuilt penalization functions: \textbf{LASSO, RIDGE} regression 
      \item \texttt{\{Boruta,caret,glmnet\}}
\end{itemize}

## Using Correlation

```{r out.width="70%", out.height="60%", size="tiny", fig.align="center", warning=F, message=F}
library(GGally)
library(dplyr)
set.seed(1234)

# Plotting the feature correlations.
iris %>% select(-Species) %>% ggpairs()
```

## Using PCR

```{r out.width="70%", out.height="60%", size="tiny", fig.align="center", warning=F, message=F}
library(dplyr)
set.seed(1234)

# Plotting the feature importance.
pcomp_df <- iris %>% 
  select(-Species) %>% prcomp(scale. = T, center = T) %>%
  plot(type="l", main = "Principle Components")
```


## Example of \texttt{\{Boruta\}}

```{r out.width="70%", out.height="60%", size="tiny", fig.align="center"}
library(Boruta)
set.seed(1234)

# Boruta is a feature selection algorithm based on the random forests algorithm.
boruta_df <- Boruta(Species ~ ., data=iris, doTrace=0)

# Plotting the feature importance.
plot(boruta_df, xlab="Features", main="Variable Importance")
```  

## Example of \texttt{\{caret\}}

```{r out.width="70%", out.height="60%", size="tiny",fig.align="center",  warning=F, message=F}
library(caret)
set.seed(1234)

# Build a decision tree model using rpart (Recursive Partitioning And Regression Trees)
rPart_df <- train(Species ~ ., data=iris, method="rpart")
rPart_imp <- varImp(rPart_df)

# Plotting the feature importance.
plot(rPart_imp, top = 3, main='Variable Importance', ylab = "Features")
```  

## Explore (E)
- Examination of data, features, and their characteristics.
 \vspace{2mm}
\begin{itemize}
      \item Data types: numerical, ordinal, and nominal data.
      \item Summary statistics.
      \item Feature distributions.
      \item Feature correlations (positive, negative).
      \item Classification: class distribution (\textbf{Class Imbalance?})
\end{itemize}
 \vspace{2mm}
- Invest your time more on the data exploration process.
\begin{itemize}
      \item Frequency distribution: \textbf{Histograms}
      \item Outlier detection: \textbf{Box plots}
      \item Feature correlation analysis: \textbf{Scatter plots}
      \item Time series analysis: \textbf{Trend and Seasonal plots}
\end{itemize}


## Tools available for Exploration
\begin{figure}
  \includegraphics[width=.9\textwidth,height=.9\textheight,keepaspectratio]{images/data_exploration.png}
  \caption{Plots available from \texttt{\{ggplot2\}}%
    \footnote{%
     \tiny{Image source: https://www.pinterest.com.au/pin/281686151677624808/}
    }%
  }
\end{figure}

## Seasonal plot from \texttt{\{feasts\}}
\begin{figure}
  \includegraphics[width=.55\textwidth,height=.55\textheight,keepaspectratio]{images/daily_seasonal.png}
  \caption{The presence of multiple seasonal cycles%
    \footnote{%
     \tiny{Github repo: https://github.com/kasungayan/Meldatathon2020} 
    }%
  }
\end{figure}

## Model Development (MD) 
- Model parameter estimation, Hyper-parameter tuning.
\vspace{2mm}
\begin{figure}
  \includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{images/model_validation.PNG}
  \caption{Model Training and Validation%
    \footnote{%
     \tiny{Image source: https://towardsdatascience.com/5-steps-of-a-data-science-project-lifecycle-26c50372b492} 
    }%
  }
\end{figure}

## Model Development Techniques
- **Supervised Learning** or **Unsupervised Learning**
 \vspace{2mm}
    \begin{itemize}
      \item Regression: Linear-regression, Support Vector Machine (SVM), Lasso-Regression
      \item Classification: Naive Bayes, Random Forest, K-Nearest Neighbors
      \item Clustering: K-Means, Fuzzy C-Means, Self Organising Maps (SOM) 
    \end{itemize}
\vspace{2mm}
- Neural Networks: Multilayer Perceptron (**MLP**), Recurrent Neural Network (**RNN**), Convolutional Neural Network (**CNN**)
\vspace{2mm}
- Different applications: **Spam Detection**, **Market Segmentation**, **Image Classification**, **Time Series Forecasting**, **Language Translation**, **Voice Recognition**

## Naive Bayes classifier

```{r size="tiny",fig.align="center", warning=F}
library(mlbench) # multiple benchmark datasets for different machine learning tasks.
library(caret) # multiple inbuilt regression and classification algorithms.
library(rsample) # data splitting.
library(dplyr)

data(BreastCancer)
set.seed(1234)

# Splitting the data into train and test sets.
df_BreastCancer_split <- initial_split(BreastCancer, prop = .7)
# Similar splitting using caTools
# df_BreastCancer_split <- sample.split(BreastCancer,SplitRatio = 0.7)

df_BreastCancer_train <- training(df_BreastCancer_split)
df_BreastCancer_test  <- testing(df_BreastCancer_split)

# Checking for class distribution.
table(df_BreastCancer_train$Class) %>% prop.table()
table(df_BreastCancer_test$Class) %>% prop.table()
```  

## Naive Bayes classifier Cont.
```{r size="tiny",fig.align="center", warning=F}
# create feature and class attributes.
features <- setdiff(names(df_BreastCancer_train), "Class")
train_features <- df_BreastCancer_train[, features]
train_class <- df_BreastCancer_train$Class

# Define a 10-fold cross validation procedure.
train_control <- trainControl(method = "cv", number = 10)

# train the naive bayes model
model_nb1 <- train(x = train_features, y = train_class, method = "nb", trControl = train_control)
#Show the confusion matrix.
confusionMatrix(model_nb1)
```  

## Naive Bayes classifier Cont.
```{r size="tiny",fig.align="center", warning=F}

# hyper-parameter grid
hyper_search_grid <- expand.grid(usekernel = c(TRUE, FALSE), fL = 0:5, adjust = seq(0, 5, by = 1))

# train the naive bayes model using a hyper-parameter grid.
model_nb2 <- train(x = train_features, y = train_class, method = "nb", 
                   trControl = train_control, tuneGrid = hyper_search_grid)

# Printing best models.
# model_nb2$results %>% 
#  arrange(desc(Accuracy)) %>% head(4)

# results for best model
confusionMatrix(model_nb2)
```  

## Naive Bayes results on the testset
```{r size="tiny",fig.align="center", warning=F}
# Applying the best model to unseen (test) dataset.
prediction_test <- predict(model_nb2, newdata = df_BreastCancer_test)
# Printing the confusing matrix on the test set.
confusionMatrix(prediction_test, df_BreastCancer_test$Class)
```  

## Generating the ROC curve
```{r out.width="70%", out.height="60%", size="tiny",fig.align="center", warning=F}
library(caTools) #to generate ROC curves

prob_results <- predict(model_nb2, df_BreastCancer_test, type = "prob")

# Generating the ROC curve for the test set.
caTools::colAUC(prob_results, df_BreastCancer_test[["Class"]], plotROC = TRUE)

```  

## Decision tree
```{r size="tiny",fig.align="center", warning=F}
library(caret)
library(rpart)

set.seed(1234)

# Remove incomplete records.
df_BreastCancer_train <- df_BreastCancer_train[complete.cases(df_BreastCancer_train), ]

# Define a 10-fold cross validation procedure.
train_control <- trainControl(method = "cv", number = 10)

# train the naive bayes model.
model_dt <- train(Class ~ ., data=df_BreastCancer_train, method = "rpart", trControl = train_control)
#Show the confusion matrix.
confusionMatrix(model_dt)
```  

## Visualising the decision tree
```{r out.width="80%", out.height="70%", size="tiny",fig.align="center", warning=F, message=F}
library(rattle)
set.seed(1234)
# Generating the decision tree.
# You can also separately use the rpart and rpart.plot to reproduce this.

fancyRpartPlot(model_dt$finalModel)

``` 

## Regression using \texttt{\{caret\}}
```{r out.width="80%", out.height="70%", size="tiny",fig.align="center", warning=F, message=F}
library(caret)
library(dplyr)
library(rsample)

set.seed(1234)

# Reading the data.
df_households <- read.csv("data/realestate.csv")
df_households_filtered <- df_households[, c(-1,-2)]

# Splitting the data.
colnames(df_households_filtered)[6] <- "price"
data_split <- initial_split(df_households_filtered, prop = .7) 

train <- training(data_split)
test <- testing(data_split)

# Defining the feature variables.
x <- model.matrix(price~., train)
# Defining the class variable.
y <- train$price
``` 

## Regression using \texttt{\{caret\}} (2)
```{r out.width="80%", out.height="70%", size="tiny",fig.align="center", warning=F, message=F}
# Linear regression model training.
model <- train(price ~ ., train, method = "lm",trControl = trainControl(method = "cv",
                                        number = 10))

# Model summary.
print(model)
``` 

## Regression using \texttt{\{glmnet\}}
```{r, size="tiny",fig.align="center", warning=F, message=F}
library(glmnet)
library(dplyr)
library(rsample)

set.seed(1234)

# Reading the data.
df_households <- read.csv("data/realestate.csv")
df_households_filtered <- df_households[, c(-1,-2)]

# Splitting the data.
colnames(df_households_filtered)[6] <- "price"
data_split <- initial_split(df_households_filtered, prop = .7) 
train <- training(data_split)
test <- testing(data_split)

# Defining the feature variables.
x <- model.matrix(price~., train)
# Defining the class variable.
y <- train$price

# Use cross validation to determine the optimal lambda.
cv <- cv.glmnet(x, y, alpha = 1)
# Fit the final model (lasso regression) on the training data
best_model <- glmnet(x, y, alpha = 1, lambda = cv$lambda.min)
``` 

## Regression using \texttt{\{glmnet\}} (2)
```{r, size="tiny",fig.align="center", warning=F, message=F}
# Evaluating the model on the test data.
x.test <- model.matrix(price~ ., test)
price_predictions <- best_model %>% predict(x.test) %>% as.numeric()

# Model performance summary.
data.frame(
  RMSE = RMSE(price_predictions, test$price),
  Rsquare = R2(price_predictions, test$price)
)
``` 


## Unsupervised Learning
- Learning patterns from unlabbled data.
 \vspace{2mm}
- Clustering ?
 \vspace{2mm}
    \begin{itemize}
      \item K means: Computationally efficient, \textbf{Optimal K ?, Outliers?}
      \item Elbow and Silhouette methods to determine the optimal K
      \item \textbf{DBSCAN}: No restrictions on the cluster shapes
      \item Feature are categorical ? \textbf{Partitioning Around Medoids (PAM)}
      \item Hierarchical clustering: \textbf{cluster dendrogram}
    \end{itemize}
\vspace{2mm}
- Auto-Encoders ?


## Kmeans clustering
```{r out.width="60%", out.height="50%",size="tiny",fig.align="center", warning=F, message=F}
library(factoextra)

set.seed(1234)

# Removing the categorical class.
df <- iris[, -5]
# Applying kmeans algorithm with k = 3.
cluster_output <- kmeans(df, centers = 3, nstart = 25)
# Illustrating the cluster distribution.
fviz_cluster(cluster_output, data = df)
``` 

## Optimal K
```{r out.width="70%", out.height="60%",size="tiny",fig.align="center", warning=F, message=F}
library(gridExtra)

set.seed(1234)
# Different methods to determine the optimal K.
elbow <- fviz_nbclust(df, kmeans, method = "wss")
silhouette <- fviz_nbclust(df, kmeans, method = "silhouette")

grid.arrange(elbow, silhouette, nrow = 1)
``` 

## Clustering using DBSCAN
```{r out.width="70%", out.height="60%",size="tiny",fig.align="center", warning=F, message=F}
library(factoextra)
library(fpc)

set.seed(1234)
df <- iris[, -5]

# Using DBSCAN algorithm without setting the K.
db_output <- fpc::dbscan(df, eps = 0.95, MinPts = 5)
fviz_cluster(db_output, df, stand = FALSE, frame = FALSE, geom = "point")
``` 

## Neural Networks
- Strong computational systems that mimic human brain.
- Hold the universal approximation property.
- Backpropagation algorithm for training.
- State-of-the-art for many prediction/classification applications.
- Different variants of neural networks.
    \begin{itemize}
      \item Multi-Layer Perceptrons (MLP): \texttt{\{nnet,neuralnet\}}
      \item Recurrent Neural Networks (RNN): \texttt{\{rnn,RSNNS\}}
      \item Convolutional neural networks (CNNs)
    \end{itemize}
- \textbf{Tensorflow, Keras, Torch} APIs in R

## Feedforward Neural Networks using \texttt{\{nnet\}}
```{r size="tiny",fig.align="center", warning=F, message=F, results='hide'}
library(caret)

data(diamonds)
set.seed(1234)

# Preprocessing.
# Use model.matrix for categorical variables (one-hot encoding)
df_diamonds <- diamonds[, c(-2, -3, -4)]

# Data splitting.
data_split <- initial_split(df_diamonds, prop = .7)
train <- training(data_split)
test  <- testing(data_split)

# You can directly use the nnet function from the nnet packaage
# mlp_fit <- nnet(price ~ ., train, size = 2, rang = 0.1, decay = 5e-4, maxit = 200)

# You can use the nnet as the method in the caret function.
mlp_fit <- train(price ~ ., train, method = "nnet",trControl = trainControl(method = "cv",
                                        number = 10))
#print(mlp_fit)
``` 

## Deep Neural Netowrks using \texttt{\{keras\}}
```{r size="tiny",fig.align="center", warning=F, message=F, results='hide'}
library(keras)
library(rsample)
library(ggplot2)
library(dplyr)

data(diamonds)
set.seed(1234)

# The initial preprocessing.
# Use model.matrix for categorical variables (one-hot encoding)
df_diamonds <- diamonds[, c(-2, -3, -4)]

# Data splitting.
data_split <- initial_split(df_diamonds, prop = .7)
train <- training(data_split)
test  <- testing(data_split)

features <- setdiff(names(train), "price")
train_features <- train[, features]
train_class <- train$price
``` 

## Model definition
```{r size="tiny",fig.align="center", warning=F, message=F}
# Defining the model
MLP_model <- keras_model_sequential() %>%
  # The overall neural network architecture
  layer_dense(units = 10, activation = "relu", input_shape = ncol(train_features)) %>%
  layer_batch_normalization() %>% layer_dense(units = 5, activation = "relu") %>%
  layer_batch_normalization() %>% layer_dropout(rate = 0.2) %>% layer_dense(units = 1) %>%
  # Defining the optimization algorithm and loss function. 
  # Use binary and multi-categorical cross entropy for classification.
  compile(optimizer = optimizer_rmsprop(),loss = "mse",metrics = c("mae"))

# print the summary of the model
summary(MLP_model)
``` 

## Model Training
```{r out.width="60%", out.height="50%", size="tiny",fig.align="center", warning=F, message=F, results='hide'}

# train the model
learn <- MLP_model %>% fit(
  x = as.matrix(train_features),
  y = as.matrix(train_class),
  epochs = 15,
  batch_size = 32,
  validation_split = .2,
  verbose = TRUE
)

plot(learn)
``` 

## Model Results
```{r out.width="60%", out.height="50%", size="tiny",fig.align="center", warning=F, message=F, results='hide'}
# Defining the real test features and actual test output.
test_features <- test[, features]
test_class <- test$price

# Generating predictions for the first 5 records in the test set.
model_predictions <- MLP_model %>% predict(as.matrix(test_features[1:5,]))
model_predictions[model_predictions < 0] <- 0

# Generating error summary statistic.
model_results <- MLP_model %>% evaluate(as.matrix(test_features), as.matrix(test_class))
``` 

## Hyper-parameter optimization
- Determining the optimal hyper-parameters for a machine learning algorithm.
 \vspace{2mm}
- Important for models with large number of hyper-parameters (neural networks)
    \begin{itemize}
      \item Random search
      \item Grid search
      \item Bayesian optimization: \texttt{\{rBayesianOptimization,mlrMBO\}}
      \item Genetic algorithm: \texttt{\{GA\}}
    \end{itemize}
 \vspace{2mm}
- An intelligent hyper-parameter search ?

## Bayesian optimization using \texttt{\{rBayesianOptimization\}}
```{r size="tiny",fig.align="center", warning=F, message=F, results='hide'}
library(caret)
library(rBayesianOptimization)
set.seed(1234)

df_households <- read.csv("data/realestate.csv")
df_households_filtered <- df_households[, c(-1,-2)]
colnames(df_households_filtered)[6] <- "price"
# Data splitting.
data_split <- initial_split(df_households_filtered, prop = .7)
train <- training(data_split)
test  <- testing(data_split)
# Define a 10-fold cross validation procedure.
train_control <- trainControl(method = "cv", number = 5)
# Defining the fit of the SVM model
svm_model_fit <- function(logC, logSigma) {
  model <- train(price ~ ., data = train, method = "svmRadial", metric = "RMSE",
                  trControl = train_control,tuneGrid = data.frame(C = exp(logC), sigma = exp(logSigma)))
  list(Score = -getTrainPerf(mod)[, "TrainRMSE"], Pred = 0)
}
## Define the bounds and search for hyper-parameters.
lower_bounds <- c(logC = -8, logSigma = -10)
upper_bounds <- c(logC = 15, logSigma = -0.65)
svm_bounds <- list(logC = c(lower_bounds[1], upper_bounds[1]), logSigma = c(lower_bounds[2], upper_bounds[2]))
# svm_ba_search <- BayesianOptimization(svm_model_fit,
#    bounds = svm_bounds,init_grid_dt = NULL, init_points = 0,n_iter = 4, verbose = TRUE)
```

## Light Gradient Boosting Machine (LightGBM)
- Gradient boosting framework that uses tree based learning algorithms.
 \vspace{2mm}
- Used for both classification and regression tasks.
 \vspace{2mm}
- Leading algorithm in many machine learning competitions (Kaggle).
\vspace{2mm}
    \begin{itemize}
      \item Faster training speed and higher efficiency
      \item Lower memory usage
      \item Highly scalable
      \item Highly parallelizable
      \item Better accuracy than any other boosting algorithms
\end{itemize}

## LightGBM using \texttt{\{lightgbm\}}
```{r size="tiny",fig.align="center", warning=F, message=F, results='hide'}
library(lightgbm)
set.seed(1234)

data("iris")
df <- iris

# Data preparation.
df$Species <- as.numeric(as.factor(df$Species)) - 1
# Data splitting.
data_split <- initial_split(df, prop = .7)
train <- training(data_split)
test  <- testing(data_split)

# Transforming to lightgbm data input format.
dtrain <- lgb.Dataset(data = as.matrix(train[1:4]), label = as.matrix(train[,5]))
dtest <- lgb.Dataset.create.valid(dtrain, data = as.matrix(test[, 1:4]), label = as.matrix(test[, 5]))
valids <- list(test = dtest)
# Defining the objective function for a multi-class problem.
params <- list(objective = "multiclass", metric = "multi_error", num_class = 3)

# Training lightgbm model.
lightgbm_model <- lgb.train(params, dtrain,100, valids, min_data = 1,learning_rate = 1.0
    ,early_stopping_rounds = 10)

# Lightgbm predicts all probabilities for the 3 classes; use argmax() to get the classified class.
lightgbm_predictions <- predict(lightgbm_model, as.matrix(test[, 1:4]), reshape = TRUE)
```

## Model Explainability
- Majority of the machine learning models are black-box.
 \vspace{2mm}
- Interpreting and explaining model predictions.
 \vspace{2mm}
- Explainable machine learning (GDPR law: Right explanation)
\vspace{2mm}
    \begin{itemize}
      \item Global and Local Interpretable models
      \item \textbf{LIME} (Local Interpretable Model-Agnostic Explanations): \texttt{\{limer\}}
      \item \textbf{SHAP} (SHapley Additive exPlanations): \texttt{\{shapr\}}
      \item \textbf{LORE} (Rule-based Explanations)
      \item \textbf{Anchor}
\end{itemize}

## Model Explainability using \texttt{\{shapr\}}
```{r size="tiny",fig.align="center", warning=F, message=F, results='hide'}
library(xgboost)
library(shapr)
set.seed(1234)

# We are explaining the prediction of household prices.
df_households <- read.csv("data/realestate.csv")
df_households_filtered <- df_households[, c(-1,-2)]
colnames(df_households_filtered)[6] <- "price"

data_split <- initial_split(df_households_filtered, prop = .7)
train <- training(data_split)
test  <- testing(data_split)

features <- setdiff(names(train), "price")
train_features <- as.matrix(train[, features])
train_class <- as.matrix(train$price)
test_features <- as.matrix(test[, features])

# Fitting a basic xgboost model to the training data
xgboost_model <- xgboost(data=train_features, label=train_class, nround=10, verbose = FALSE)

# Prepare the data for explanation
model_explainer <- shapr(train_features, xgboost_model)
# Expected value without the prediction.
expected_price <- mean(train_class)
```


## Model Explainability (2)
```{r out.width="80%", out.height="70%", size="tiny",fig.align="center", warning=F, message=F, results='hide'}

explanation <- explain(test_features, approach = "empirical", 
                       explainer = model_explainer, prediction_zero = expected_price)

# Plot the resulting explanations for observations 1 and 6
plot(explanation, plot_phi0 = FALSE, index_x_test = c(2, 4, 8, 16))

```

## References

\tiny 
[1] Hadley Wickham (2021). rvest: Easily Harvest (Scrape) Web Pages. R package version 1.0.0.
  https://CRAN.R-project.org/package=rvest
  
[2] Mark van der Loo (2021). simputation: Simple Imputation. R package version 0.2.6. https://CRAN.R-project.org/package=simputation
  
[3]  https://cran.r-project.org/web/packages/simputation/vignettes/intro.html

[4] https://www.analyticsvidhya.com/blog/2016/12/introduction-to-feature-selection-methods-with-an-example-or-how-to-select-the-right-variables/

[5] https://www.kaggle.com/zurfer/how-to-use-principle-component-analysis-in-r

[6] Miron B. Kursa, Witold R. Rudnicki (2010). Feature Selection with the Boruta Package. Journal of Statistical Software,
  36(11), 1-13. URL http://www.jstatsoft.org/v36/i11/
  
[7] Max Kuhn (2020). caret: Classification and Regression Training. R package version 6.0-86.
  https://CRAN.R-project.org/package=caret
  
[8] Barret Schloerke, Di Cook, Joseph Larmarange, Francois Briatte, Moritz Marbach, Edwin Thoen, Amos Elberg and Jason
  Crowley (2021). GGally: Extension to 'ggplot2'. R package version 2.1.1. https://CRAN.R-project.org/package=GGally
  
[9] Mitchell O'Hara-Wild, Rob Hyndman and Earo Wang (2021). feasts: Feature Extraction and Statistics for Time Series. R
  package version 0.1.7. https://CRAN.R-project.org/package=feasts

[10] https://rpubs.com/maulikpatel/229337

## References 

\tiny 
[11] Julia Silge, Fanny Chow, Max Kuhn and Hadley Wickham (2021). rsample: General Resampling Infrastructure. R package
  version 0.0.9. https://CRAN.R-project.org/package=rsample
  
[12] https://uc-r.github.io/naive_bayes

[13] Jarek Tuszynski (2021). caTools: Tools: Moving Window Statistics, GIF, Base64, ROC AUC, etc. R package version 1.18.1.
  https://CRAN.R-project.org/package=caTools

[14] Terry Therneau and Beth Atkinson (2019). rpart: Recursive Partitioning and Regression Trees. R package version 4.1-15.
  https://CRAN.R-project.org/package=rpart
  
  
